Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-05-31 code-tests.tex

Ensuring that a complex piece of software is behaving correctly is a
non-trivial task.  While there are a range of techniques that one can
apply to ensure correctness, the \enzo\ code uses two primary methods:
a suite of test problems that can be compared to previous versions of
the code, to ensure that \enzo\ is running correctly on a new
computer, with a new compiler, or after substantial changes have been
made to the code base; and by direct comparisons to other
astrophysical fluid dynamics codes.  We describe the \enzo\ test
methodology in Section~\ref{sec.tests.suite}, code comparisons in
Section~\ref{sec.tests.compare}, and show a small set of
representative test problems in Section~\ref{sec.tests.problems}.  We
further note that the \enzo\ test suite
(Section~\ref{sec.tests.suite}) contains hundreds of test problems, as
well as the ability to compare to a ``gold standard'' solution, and
thus all of the tests shown here are easily reproducible by the
reader.

>>>>

Ensuring that a complex piece of software is behaving correctly is a
non-trivial task.  While there are a range of techniques that one can
apply to ensure correctness, the \enzo\ code uses two primary methods:
a suite of test problems that can be compared to previous versions of
the code, used to ensure that \enzo\ is running correctly on a new
computer, with a new compiler, or after substantial changes have been
made to the code base; and by direct comparisons to other
astrophysical fluid dynamics codes.  We describe the \enzo\ test
methodology in Section~\ref{sec.tests.suite}, code comparisons in
Section~\ref{sec.tests.compare}, and show a small set of
representative test problems in Section~\ref{sec.tests.problems}.  We
further note that the \enzo\ test suite
(Section~\ref{sec.tests.suite}) contains hundreds of test problems, as
well as the ability to compare to a ``gold standard'' solution, and
thus all of the tests shown here are easily reproducible by the
reader.

### 2013-05-31 tests-doublemach.tex

The double mach reflection test is a classic two-dimensional test of
hydrodynamic algorithms originally described in
\citet{1984JCoPh..54..115W} \citep[and more recently
in][]{2008ApJS..178..137S}.  In this problem, shown in
Figure~\ref{fig.doublemach}, a shock is injected at an angle to a
reflecting surface (the -y boundary), and a jet appears along the
reflecting surface.  The ideal solution is self-similar, and the
appearance of this solution is highly sensitive to numerical
diffusion.  If numerical noise is present, a Kelvin-Helmholz
instability develops along this jet and breaks the self-similarity.

In the test problem shown in Figure~\ref{fig.doublemach}, a 2D
simulation with $960 \times 240$ cells was created with a domain of $x
= [0, 4]$ and $y = [0, 1]$.  We use an ideal gas equation of state of
$\gamma = 1.4$, a pre-shock density of 1.4, and a pre-shock specific
internal energy of $2.5/1.4$ (all in arbitrary units).  A Mach 10
shock is initialized with a shock normal that is 30$^\circ$ from the
x-axis and an initial position on the lower boundary of x$ = 1/6$.
The lower y boundary and right x boundary are reflecting; the left x
boundary is inflowing, and the upper y boundary has a time-dependent
boundary condition that allows the shock to propagate into the domain
as if it extends to infinity.  The simulation starts at t$ = 0$ and
runs until t$ = 0.205$ (arbitrary units), at which point the rightmost
extent of the shock should be at roughly x$ = 3$.  In this simulation we use
the direct Eulerian implementation of the piecewise parabolic hydrodynamic
method with the diffusion, flattening, and shock steepening all enabled.

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/DoubleMachTest.eps}
\caption{Density field at the final time in the Double Mach test.  A
Mach 10 shock is injected into the domain with a shock normal that is
30$^\circ$ from the x-axis with a time-dependent +y boundary condition
that mimics a shock of infinite length.  The solution is self-similar,
with the jet and whorls at $x \simeq 2.5-3.0$ in this figure being very sensitive
to numerical diffusion.  Our results compare favorably to the
higher-order images from \citet{1984JCoPh..54..115W}.}
\label{fig.doublemach}
\end{center}
\end{figure}

>>>>

The double mach reflection test is a classic two-dimensional test of
hydrodynamic algorithms originally described in
\citet{1984JCoPh..54..115W} \citep[and more recently
in][]{2008ApJS..178..137S}.  In this problem, shown in
Figure~\ref{fig.doublemach}, a shock is injected at an angle to a
reflecting surface (the -y-boundary), and a jet appears along the
reflecting surface.  The ideal solution is self-similar, and the
appearance of this solution is highly sensitive to numerical
diffusion.  If numerical noise is present, a Kelvin-Helmholz
instability develops along this jet and breaks the self-similarity.

In the test problem shown in Figure~\ref{fig.doublemach}, a 2D
simulation with $960 \times 240$ cells was created with a domain of $x
= [0, 4]$ and $y = [0, 1]$.  We use an ideal gas equation of state of
$\gamma = 1.4$, a pre-shock density of 1.4, and a pre-shock specific
internal energy of $2.5/1.4$ (all in arbitrary units).  A Mach 10
shock is initialized with a shock normal that is 30$^\circ$ from the
x-axis and an initial position on the lower boundary of x$ = 1/6$.
The lower y-boundary and right x-boundary are reflecting; the left x
boundary is inflowing, and the upper y-boundary has a time-dependent
boundary condition that allows the shock to propagate into the domain
as if it extends to infinity.  The simulation starts at t$ = 0$ and
runs until t$ = 0.205$ (arbitrary units), at which point the rightmost
extent of the shock should be at roughly x$ = 3$.  In this simulation we use
the direct Eulerian implementation of the piecewise parabolic hydrodynamic
method with the diffusion, flattening, and shock steepening all enabled.

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/DoubleMachTest.eps}
\caption{Density field at the final time in the Double Mach test.  A
Mach 10 shock is injected into the domain with a shock normal that is
30$^\circ$ from the x-axis with a time-dependent +y-boundary condition
that mimics a shock of infinite length.  The solution is self-similar,
with the jet and whorls at $x \simeq 2.5-3.0$ in this figure being very sensitive
to numerical diffusion.  Our results compare favorably to the
higher-order images from \citet{1984JCoPh..54..115W}.}
\label{fig.doublemach}
\end{center}
\end{figure}

### 2013-05-31 tests-shockpool.tex

The next problem is similar to the last one but instead examines how a shock with a Mach number of 2 passes in and out of a static refined region (again defined from $z=0.25$ to 0.75).  The density and pressure in the domain is initially set equal to 1.0 with zero velocity.  At $t=0$ the left boundary is set with the density, pressure and velocity appropriate for $\mathcal{M}=2$ shock wave.

@@

For Zeus (middle row of Figure~\ref{fig.shockpool}, the shock is broadened because of artificial viscosity and there are slightly more post-shock oscillations, although again quite mild.  The impact of entering and exiting the refined region is somewhat larger than in PPM; however, the most noticeable difference is the incorrect position of the shock front. This is due to the fact that the scheme is not energy-conserving (see also the Sedov problem, below), and has very little to do with the passage through the refined region.

>>>>

The next problem is similar to the last one but instead examines how a shock with a Mach number of 2 passes in and out of a static refined region (again defined from $z=0.25$ to 0.75).  The density and pressure in the domain is initially set equal to 1.0 with zero velocity.  At $t=0$ the left boundary is set with the density, pressure and velocity appropriate for a $\mathcal{M}=2$ shock wave.

@@

For Zeus (middle row of Figure~\ref{fig.shockpool}), the shock is broadened because of artificial viscosity and there are slightly more post-shock oscillations, although again quite mild.  The impact of entering and exiting the refined region is somewhat larger than in PPM; however, the most noticeable difference is the incorrect position of the shock front. This is due to the fact that the scheme is not energy-conserving (see also the Sedov problem, below), and has very little to do with the passage through the refined region.
